---
title: "p8105_hw1_yw4817"
output: github_document
date: "2026-09-26"
---
#Problem 1

1) Write a short description of the penguins dataset (not the penguins_raw dataset) using inline R code. In your discussion, please include:
-the data in this dataset, including names / values of important variables
-the size of the dataset (using nrow and ncol)
-the mean flipper length

Answer:
```{r}
data("penguins", package = "palmerpenguins")
```
The `penguins` contains 8 variables, which are species, island, bill length, bill depth, flipper lenth, body mass, sex, and year. Important variables are species, island, and sex. Species include `r paste(unique(penguins$species),collapse=", ")`; island includes `r paste(unique(penguins$island),collapse=", ")`; sex includes `r paste(unique(penguins$sex),collapse=", ")`.
The "penguins" dataset contains `r nrow(penguins)` observations and `r ncol(penguins)` variables.
The mean of flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)`.

2) Make a scatterplot of flipper_length_mm (y) vs bill_length_mm (x); color points using the species variable (adding color = ... inside of aes in your ggplot code should help). Export your first scatterplot to your project directory using ggsave.

Answer:
```{r,message = FALSE}
library(tidyverse)
```

```{r}
ggplot(
  data = penguins,
  aes(
    x = bill_length_mm,
    y = flipper_length_mm,
    color = species
    )
  ) + geom_point()

ggsave("penguins_scatterplot.pdf")
```
#Problem 2

1) Create a data frame comprised of:
-a random sample of size 10 from a standard Normal distribution
-a logical vector indicating whether elements of the sample are greater than 0
-a character vector of length 10
-a factor vector of length 10, with 3 different factor “levels”

Answer:
```{r}
sample = rnorm(10)
sample_positive = sample > 0
char_vector = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j")
factor_vector = factor(c("A", "B", "C", "A", "B", "C", "A", "B", "C", "A"))
```
Data Frame:
```{r}
p2_df = data.frame(
  sample = sample,
  sample_positive = sample_positive,
  char_vector = char_vector,
  factor_vector = factor_vector)
```

2) Try to take the mean of each variable in your dataframe. What works and what doesn’t?

Answer:
```{r}
mean(pull(p2_df, sample))
mean(pull(p2_df, sample_positive))
mean(pull(p2_df, char_vector))
mean(pull(p2_df, factor_vector))
```
The mean works for numerical and logical variables and does not work for character and factor variables. This is because numerical values are numbers, and R treats TRUE as 1 and FALSE as 0 for logical variables. Characters and factors are in text or represent categories, which cannot be calculated directly.

3) Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). What happens, and why? Does this help explain what happens when you try to take the mean?

Answer:
```{r, results='hide', warning = FALSE}
as.numeric(pull(p2_df, sample_positive))
as.numeric(pull(p2_df, char_vector))
as.numeric(pull(p2_df, factor_vector))
```
The logical variables are converted to numerical form as 1 and 0 after applying as.numeric(). Factor variables are also converted to numbers, while character variables return NA because they cannot be converted to numbers. This is because factor variables return the numeric codes corresponding to their factor levels when applying as.numeric(). This helps explain why the mean works for the logical variable but not for the character and factor variables.
